### 2021-11-13 manuscript.Rmd
coding style

keywords          : "sequential analysis, hypothesis testing, efficiency, estimation, bias"
wordcount         : "11,368"

@@

r_refs(file = "r-refs.bib") #update later
default_plot_hook = knitr::knit_hooks$get('plot')

@@

gs  <- function(proc, return = "n") {
  bs = ifelse(proc == "asOF", "bsOF", "bsP") #specify beta-spending function
  if(proc == "asOF") {rates = c(0.50, 0.75, 1)} else{rates = c(1/3, 2/3, 1)} #specify information rates for OBF vs Pocock

@@

  # Get parameters
  parameters <- getSampleSizeMeans(design = design, groups = 2, alternative = 0.5)
  n_gs = ceiling(c(parameters$numberOfSubjects[1], diff(parameters$numberOfSubjects))/2) # n per look per group
  alpha <- parameters$criticalValuesPValueScale
  futility <- parameters$futilityBoundsPValueScale %>% round(2)

  if(return == "alpha") {return(alpha %>% signif(2) %>% as.character() %>% sub(".", "", .,))}
  if(return == "n") {return(n_gs)}
  if(return == "futility") {return(futility %>% as.character() %>% sub(".", "", .,))}
}

@@

For the group-sequential designs, we can see that the available median-unbiased estimators provide a great improvement over the naive (Cohen's *d*) estimator. For both the Pocock-like and the O'Brien-Fleming-like designs, the bias-adjusted effect size estimate is very close to the true population effect size: *d* = 0.49 compared to $\delta$ = 0.50. For the SBF procedure, we see that the using the mean of the posterior distribution (which shrinks the effect size estimate in earlier terminations) also provides a slight improvement in parameter estimation (by reducing the estimated effect size from *d* = 0.57 to *d* = 0.55).  The bias in the median effect size estimate of the ISP, however, is severe (*d* = 0.66; an exaggeration ratio of 1.33). This biased estimate remains uncorrected due to the absence of a bias-adjusted estimator accompanying the ISP. A we can see, the distribution of effect size estimates stemming from the ISP is bimodal.

@@

```{r include = F}
#df <- read.csv("simulations/data.csv")

df = df %>% 
  mutate(
    name = case_when(
      proc == "Fixed" ~ "Fixed",
      proc == "Bayes" ~ "Bayes",
      proc == "ISP" ~ "ISP",
      proc == "asP" ~ "Pocock",
      proc == "asOF"~ "O'Brien-Fleming"
      ),
    name = fct_relevel(name, "Fixed", "Pocock", "O'Brien-Fleming", "Bayes", "ISP")
    )

t1 = df %>% filter(d_forpower == d_actual) %>%
  group_by(name) %>% 
  summarize(true_positive = mean(decision == "reject.null") %>% round(2),
            inconclusive = mean(decision == "inconclusive") %>% round(2),
            false_negative = mean(decision == "reject.alt") %>% round(2))

t2 = df %>% filter(d_actual == 0) %>% 
  group_by(name) %>% 
  summarize(true_negative = mean(decision == "reject.alt") %>% round(2),
            inconclusive = mean(decision == "inconclusive") %>% round(2),
            false_positive = mean(decision == "reject.null") %>% round(2))
```

```{r, echo = FALSE}
knitr::kable(t2, 
             col.names = c("Procedure", "True Negative", "Inconclusive", "False Positive"), 
             align = "lccc")
```

@@

```{r, echo = FALSE}
knitr::kable(t1, 
             col.names = c("Procedure", "True Positive", "Inconclusive", "False Negative"), 
             align = "lccc")
```

@@

```{r, echo = FALSE}
E_n = df %>% 
  group_by(name, d_actual) %>% 
  summarize(E_n = mean(n_cumulative)) %>% 
  rename(Procedure = name)

t3 = E_n %>% mutate(E_n = E_n %>% round()) %>% 
  pivot_wider(names_from = d_actual, names_prefix = "*d* = ", values_from = E_n)

@@

median <- df_long %>% 
  group_by(name, d_actual, ES_type) %>% 
  summarize(medianES = round(median(ES), 2)) %>% 
  rename(Procedure = name) %>% 
  filter(!((ES_type == "ES_corrected") & (Procedure == "Fixed" | Procedure == "ISP"))) %>% 
  rowwise() %>% 
  mutate(
    name = case_when(
      ES_type == "ES" ~ as.character(Procedure),
      ES_type == "ES_corrected" ~ paste(Procedure, "adjusted")
      )
    ) %>% 
  ungroup() %>% 
  select(name, d_actual, medianES) %>% 
  rename(Procedure = name)

t4 <- median %>% 
  pivot_wider(
    names_from = d_actual, 
    names_prefix = "*d* = ", 
    values_from = medianES
    )

t4
```

>>>>

keywords          : "sequential analysis, hypothesis testing, efficiency, error rates, parameter estimation, sample size"
wordcount         : "11,368"

@@

r_refs(file = "r-refs.bib") #update later
default_plot_hook <- knitr::knit_hooks$get('plot')

@@

gs  <- function(proc, return = "n") {
  bs <- ifelse(proc == "asOF", "bsOF", "bsP") #specify beta-spending function
  if(proc == "asOF") {
    rates = c(0.50, 0.75, 1)
    } else{
      rates = c(1/3, 2/3, 1)
      } #specify information rates for OBF vs Pocock

@@

  # Get parameters
  parameters <- getSampleSizeMeans(design = design, groups = 2, alternative = 0.5)
  n_gs <- ceiling(c(parameters$numberOfSubjects[1], diff(parameters$numberOfSubjects))/2) # n per look per group
  alpha <- parameters$criticalValuesPValueScale
  futility <- parameters$futilityBoundsPValueScale %>% round(2)

  if(return == "alpha") {
    return(
      alpha %>% 
        signif(2) %>% 
        as.character() %>% 
        sub(".", "", .,))
    }
  if(return == "n") {return(n_gs)}
  if(return == "futility") {
    return(
      futility %>% 
        as.character() %>% 
        sub(".", "", .,))
    }
}

@@

For the group-sequential designs, we can see that the available median-unbiased estimators provide a great improvement over the naive (Cohen's *d*) estimator. For both the Pocock-like and the O'Brien-Fleming-like designs, the bias-adjusted effect size estimate is very close to the true population effect size: *d* = 0.49 compared to $\delta$ = 0.50. For the SBF procedure, we see that the using the mean of the posterior distribution (which shrinks the effect size estimate in earlier terminations) also provides a slight improvement in parameter estimation (by reducing the estimated effect size from *d* = 0.57 to *d* = 0.55).  The bias in the median effect size estimate of the ISP, however, is severe (*d* = 0.66; an exaggeration ratio of 1.33). This biased estimate remains uncorrected due to the absence of a bias-adjusted estimator accompanying the ISP. As we can see, the distribution of effect size estimates stemming from the ISP is bimodal.

@@

```{r include = F}
df <- df %>% 
  mutate(
    name = case_when(
      proc == "Fixed" ~ "Fixed",
      proc == "Bayes" ~ "Bayes",
      proc == "ISP" ~ "ISP",
      proc == "asP" ~ "Pocock",
      proc == "asOF"~ "O'Brien-Fleming"
      ),
    name = fct_relevel(name, "Fixed", "Pocock", "O'Brien-Fleming", "Bayes", "ISP")
    )

t1 <- df %>% 
  filter(d_forpower == d_actual) %>%
  group_by(name) %>% 
  summarize(
    true_positive = mean(decision == "reject.null") %>% round(2),
    inconclusive = mean(decision == "inconclusive") %>% round(2),
    false_negative = mean(decision == "reject.alt") %>% round(2)
    )

t2 <- df %>% 
  filter(d_actual == 0) %>% 
  group_by(name) %>% 
  summarize(
    true_negative = mean(decision == "reject.alt") %>% round(2),
    inconclusive = mean(decision == "inconclusive") %>% round(2),
    false_positive = mean(decision == "reject.null") %>% round(2)
    )
```

```{r, echo = FALSE}
knitr::kable(
  t2, 
  col.names = c("Procedure", "True Negative", "Inconclusive", "False Positive"), 
  align = "lccc"
  )
```

@@

```{r, echo = FALSE}
knitr::kable(
  t1, 
  col.names = c("Procedure", "True Positive", "Inconclusive", "False Negative"), 
  align = "lccc"
  )
```

@@

```{r, echo = FALSE}
E_n <- df %>% 
  group_by(name, d_actual) %>% 
  summarize(E_n = mean(n_cumulative)) %>% 
  rename(Procedure = name)

t3 <- E_n %>%
  mutate(
    E_n = E_n %>% round()
    ) %>% 
  pivot_wider(
    names_from = d_actual, 
    names_prefix = "*d* = ", 
    values_from = E_n
    )

@@

median <- df_long %>% 
  group_by(name, d_actual, ES_type) %>% 
  summarize(
    medianES = round(median(ES), 2)
    ) %>% 
  rename(Procedure = name) %>% 
  filter(!((ES_type == "ES_corrected") & (Procedure == "Fixed" | Procedure == "ISP"))) %>% 
  rowwise() %>% 
  mutate(
    name = case_when(
      ES_type == "ES" ~ as.character(Procedure),
      ES_type == "ES_corrected" ~ paste(Procedure, "adjusted")
      )
    ) %>% 
  ungroup() %>% 
  select(name, d_actual, medianES) %>% 
  rename(Procedure = name)

t4 <- median %>% 
  pivot_wider(
    names_from = d_actual, 
    names_prefix = "*d* = ", 
    values_from = medianES
    )
```
